A novell interest management algorithm for distributed simulations of MMGs

### 2008-05-01 trunk/textos/dsrt/engaoi.tex
eng

%\usepackage[brazil]{babel}   
\usepackage[latin1]{inputenc}

@@

This work is based on a virtual environment partitioned into regions, each managed by a server. The regions are contiguous, exploring the locality of the players' avatars. Thus, avatars next to each other will probably be located in the same region and, therefore, their clients will tend to be connected to the same server, so that their interaction is faster (Fig. \ref{partitions}). Situations in which two players interacting with each other are connected to different servers imply a higher bandwidth usage, as it is required some sort of negotiation between the servers to which the different players are connected, in order to maintain the states of simulation in both servers identical. Also, the hop count of each message between the clients of these players is increased, since there is one more intermediary.

@@

One issue that relates to that type of partitioning of the virtual environment concerns the borders between regions. If an avatar of a client connected to a server is close to the border of a region with another, which is managed by another server, it will be necessary to have an exchange of information between these servers. This information consists of state updates of the entities that are interacting with each other despite being located in different regions. For example, let $S_A$ be the server responsible for the region $R_A$ where is located the avatar of the client $C_A$ and let be $S_B$ the server responsible for another region, $R_B$, where is located the avatar of the client $C_B$. When the avatar of $C_A$ gets close the border with $R_B$, $S_A$ sends to $S_B$ a message notifying it about the presence of that avatar near its region's border. If the avatar of $C_B$ also gets close to the border with $R_A$ and to the avatar of $C_A$, $S_B$ sends another message, notifying $S_A$. Then, the avatars of $C_A$ and $C_B$ start exchanging state updates through their servers.

@@

This way, the details of this mechanism will not result in relevant changes to the interest management. When a player $J$, whose avatar is near the border of a region, performs actions whose results must be broadcasted to players with their avatars in other regions, the server $S$, responsible for $J$, simulates his actions, calculates the resulting state and simply sends it to the neighbor server, as if it was sending to its own clients. Similarly, when $S$ receives the resulting state of the action of a player who is connected to a neighbor server, it broadcasts the state to the clientes connected to it as if a player in its own region had executed that action.

@@

In order to provide an identical sense of the environment among the players, each one of them must maintain a copy of the entities' states, which must be the same for everyone. The simplest way to do this is to broadcast the states of all entities to all clients. The problem of this approach is that it generates a heavy traffic between the servers and clients, preventing the game to well, as the number of participants increase. To save bandwidth, both of the players, as the servers that intermediate them, a technique known as interest management is employed. This technique reduces the number of updates that the players will receive - and send, in the case of a peer-to-peer architecture.

In short, the interest management technique works as follows: for every state change of each entity, it is calculated for whom it will be important. For example, if an avatar is miles away from another in the virtual environment, it is most likely that their state changes is irrelevant to each other. Thus, it is not necessary for them to exchange information on their state. This principle - locality - is used as the main criterion in the interest management algorithms.

@@

\subsection{Field of view}
\label{angle}

A more refined method to manage the interests of the avatars is to take into account what each player can see. The area within which the player perceives state changes can be defined as a circular sector. This is similar to the circular area of interest described in the previous section, but takes into account that the player can only see objects that are located in front of his avatar.

@@

One issue to be considered, however, is that the player will not receive state updates from entities entities which are close to his avatar, but behind it. This could cause some problems. For example, if the avatar turns 180 degrees, the player might not be able to see a particular entity that should be there, needing some time to receive its state information. This happens because, despite this entity have been near the avatar, the player did not receive its information yet because it was behind him, outside his field of view. In Figure \ref{fangle} it is illustrated an area of interest that takes into account the player's view angle.

@@

The principle behind the approach proposed here is based on the fact that the more distant an entity is from the avatar in the virtual environment, the lower its update frequency for that avatar may be. Therefore, the state updates from entities which are more distant may be received with a longer interval between them. On the other hand, if an entity is very close, it is desirable that the player receives its most recent state change information as soon as possible, to view any alterations quickly.

@@

Before sending the state of an entity to a client, the last transmission time is checked. The next transmission is then scheduled to occur after a certain interval. If the relevance of that state is 1, the normal update interval will be used. If it is less than 1, the normal interval is divided by the relevance. For example, let the normal update interval of a game be 200 ms. If the avatar $A_i$, which just sent a state update packet to $A_j$, is at a distance from $A_i$ such that its relevance is 0.5, the next transmission will only happen after an interval of 200/0.5. So, the player controlling $A_j$ will only receive an update of $A_i$ within 400 ms. Despite this interval is still less than a half second, it represents a reduction of the state update frequecy of $A_i$ in 50\%. As they are at a greater distance from one another, and the interval was increased only by 200 ms, this variation will probably be imperceptible by the player who controls $A_j$.

@@

A simple form of using various update intervals based on the relevance value would be utilizing the circular area of interest. To obtain the relevance of an entity relative to an avatar, the value may be set to 1 when the entity is in the same position of the avatar and gradually reduce as it gets more distant, until it reaches 0 and stops decreasing no matter how more distant it gets. This is a way that, although simple, has shown a significant reduction in traffic between clients and servers. In Figure \ref{graded_circle}, it is illustrated how the area of interest would be, with entities' state update frequency attenuation for a given avatar.

@@

 \item Close area, whose purpose is %evitar que ocorram problemas caso o jogador faça seu avatar girar ao redor do próprio eixo muito rapidamente, além de 
 improve the game quality in the space near the avatar. Its radius is the critical distance, defined previously;
 \item Update frequency attenuation.
\end{itemize}

The resulting area of interest takes then the shape of a circular sector, which represents the player's field of view. The origin of this circular sector is also the center of a smaller circle, which defines the close area of the avatar. In the close area, all entities have a relevance value of 1, so their update frequency to that avatar is set to the maximum value. This way, the most updated game state becomes available to the player in the area around his avatar, improving the interaction with entities next to it. Even if some of them are outside the player's field of view, he will be able to see them if his avatar turns rapidly to their direction. Figure \ref{fcsmooth} illustrates the area of interest which has just been defined.

As for entities that are outside the area nearby, but still within the avatar's field of view, their relevance must be calculated. It is proposed that the relevance of each entity decline gradually in accordance with the distance between it and the avatar in question. The further they are, the less frequent will be their state updates. This is possible because even if the interval update is doubled, it will most likely still be a fraction of a second, which is hardly noticeable for a player whose avatar is located at a great distance from the entity in question. Moreover, short delays between the arrival of updates of state can easily be masked by techniques of interpolation, such as dead-reckoning \cite{smed2002rna}. The algorithm \ref{smalg} defines the operation of our interest management technique.

@@

In order to perform the simulation of the proposed algorithm, it was first necessary to create a simulated virtual environment, with avatars present on it, because the algorithm is based on locality and view angle information. The environment consists of a two-dimensional space, which corresponds to the region managed by one of the servers. There are various avatars present, whose number varies from one simulation to another. Each one of them randomly chooses a destination point in the environment and then goes there. When it reaches its destination, it stays for a random time value, which may be zero, and then chooses a new destination to go to.

Ns-2 simulator \cite{mccanne:nsn} was used to simulate and compare the interest management algorithms described in this paper. This simulator allows the user to create code of the specific application which will be simulated. In our case, it was simulated a server which should send state updates to the client responsible for one of the avatars in the region managed by that server. Based on the location of the other avatars and on the chosen interest management algorithm, the server decided which other avatars had a relevant state to the client in question.

>>>>

%\usepackage[brazil]{babel}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}

@@

This work is based on a virtual environment partitioned into regions, each managed by a different server. The regions are contiguous, exploring the locality of the players' avatars. Thus, avatars next to each other will probably be located in the same region and, therefore, their clients will tend to be connected to the same server, so that their interaction is faster (Fig. \ref{partitions}). Situations in which two players interacting with each other are connected to different servers imply a higher bandwidth usage, as it is required some sort of negotiation between the servers to which the different players are connected, in order to maintain the states of the simulation identical in both servers. Also, the hop count of each message between the clients of these players is increased, since there is one more intermediary.

@@

One issue that relates to that type of partitioning of the virtual environment concerns the borders between regions. If an avatar is close to the border between two regions, it will be necessary some exchange of information between the servers which manage them. This information consists of state updates of the entities that are interacting with each other despite being located in different regions. For example, let $S_i$ be the server responsible for the region $R_i$ where the avatar of the client $C_i$ is located and let $S_j$ be the server responsible for another region, $R_j$, where the avatar of the client $C_j$ is situated. When the avatar of $C_i$ gets close to the border with $R_j$, $S_i$ sends to $S_j$ a message notifying it about the presence of that avatar near its region's border. If the avatar of $C_j$ also gets close to the border with $R_i$ and to the avatar of $C_i$, $S_j$ sends another message, notifying $S_i$. Then, $C_i$ and $C_j$ start exchanging state updates of their avatars through the servers.

@@

This way, the details of this mechanism will not result in relevant changes to the interest management. When a player $P$, whose avatar is near the border of a region, performs actions whose results must be broadcasted to players with their avatars in other regions, the server $S$, responsible for $P$, simulates his actions, calculates the resulting state and simply sends it to the neighbor server, as if it was sending to its own clients. Similarly, when $S$ receives the resulting state of the action of a player who is connected to a neighbor server, it broadcasts the state to the clientes connected to it as if a player in its own region had executed that action.

@@

In order to provide an identical sense of the environment among the players, each one of them must maintain a local copy of the entities' states, which must be the same for everyone. The simplest way to do this is to broadcast the states of all entities to all clients. The problem of this approach is that it generates a heavy traffic between the servers and clients, preventing the game to scale well, as the number of participants increase. To save bandwidth, both of the players, as of the servers that intermediate them, a technique known as interest management is employed. This technique reduces the number of updates that the players will receive - and send, in the case of a peer-to-peer architecture.

In short, the interest management technique works as follows: for every state change of each entity, it is calculated to whom it will be important. For example, if an avatar is miles away from another one in the virtual environment, it is most likely that their state alterations are irrelevant to each other. Thus, it is not necessary for them to exchange information on their state. This principle - locality - is used as the main criterion in the interest management algorithms.

@@

\subsection{Field of view based area of interest}
\label{angle}

A more refined method to manage the players' interests is to take into account what each one of them can see. The area within which the player perceives state changes can be defined as a circular sector. This is similar to the circular area of interest described in the previous section, but considers that the player can only see objects that are located in front of his avatar.

@@

One issue to be observed, however, is that the player will not receive state updates from entities that are close to his avatar, but behind it. This could cause some problems. For example, if the avatar turns 180 degrees, the player might not be able to see a particular entity that should be there, needing some time to receive its state information. This happens because, though the entity was near the avatar, it was outside the player's field of view. In Figure \ref{fangle}, it is illustrated an area of interest that takes into account the player's view angle.

@@

The principle behind the approach proposed here is based on the fact that the more distant an entity is from the avatar in the virtual environment, the lower its update frequency for that avatar may be. Therefore, the state updates from entities which are more distant may be received with a longer interval between them. On the other hand, if an entity is very close, it is desirable that the player receives its most recent state information as soon as possible, to view any alterations quickly.

@@

Before sending the state of an entity to a client, the last transmission time is checked. The next transmission is then scheduled to occur after a certain interval. If the relevance of that state is 1, the normal update interval will be used. If it is less than 1, the normal interval is divided by the relevance. For example, let the normal update interval of a game be 200 ms. If the avatar $A_i$, which just sent a state update packet to $A_j$, is at a distance from $A_i$ such that its relevance is 0.5, the next transmission will only happen after an interval of 200 divided by 0.5. So, the player controlling $A_j$ will only receive an update of $A_i$ after 400 ms. Despite this interval is still less than a half second, it represents a reduction of the state update frequecy of $A_i$ in 50\%. As they are at a greater distance from one another, and the interval was increased only by 200 ms, this variation will probably be imperceptible by the player who controls $A_j$.

@@

A simple form of using various update intervals based on the relevance value would be utilizing the circular area of interest. To obtain the relevance of an entity relative to an avatar, the value may be set to 1 when they are in the same position and gradually reduce it as the distance between them increases, until it reaches 0 and stops decreasing no matter how further the entity goes. This is a way that, although simple, has shown a significant reduction in traffic between clients and servers. In Figure \ref{graded_circle}, it is illustrated how the area of interest would be, with entities' state update frequency attenuation for a given avatar.

@@

 \item Close area, whose purpose is %evitar que ocorram problemas caso o jogador faça seu avatar girar ao redor do próprio eixo muito rapidamente, além de 
 to improve the game quality in the space near the avatar. Its radius is the critical distance;
 \item Update frequency attenuation.
\end{itemize}

The resulting area of interest then takes the shape of a circular sector, which represents the player's field of view. The origin of this circular sector is also the center of a smaller circle, which defines the are close to the avatar. In the close area, all entities have a relevance value of 1, so their update frequency to that avatar is set to the maximum value. This way, the most up-to-date game state becomes available to the player in the area around his avatar, improving the interaction with entities next to it. Even if some of them are outside the player's field of view, he will be able to see them if his avatar turns rapidly to their direction. Figure \ref{fcsmooth} illustrates the area of interest which has just been described.

As for entities that are outside the area nearby, but still within the avatar's field of view, their relevance must be calculated. It is proposed that the relevance of each entity decline gradually as the distance between it and the avatar in question increases. The further they are, the less frequent will be their state updates. This is possible because even if the interval update is doubled, it will most likely still be a fraction of a second, which is hardly noticeable for a player whose avatar is located at a great distance from the entity in question. Moreover, short delays between the arrival of state updates can be easily masked by extrapolation techniques, such as dead-reckoning \cite{smed2002rna}. The algorithm \ref{smalg} defines the operation of our interest management technique.

@@

In order to perform the simulation of the proposed algorithm, it was first necessary to create a simulated virtual environment, with avatars present on it, since the algorithm is based on locality and view angle information. The environment consists of a two-dimensional space, which corresponds to the region managed by one of the servers. There are various avatars present, whose number varies from one simulation to another. Each one of them randomly chooses a destination point in the environment and then moves there. When it reaches its destination, it stays for a random time value, which may be zero, and then chooses a new destination to move to.

Ns-2 simulator \cite{mccanne:nsn} was used to simulate and compare the interest management algorithms described in this paper. This simulator allows the user to create code of the specific application which will be simulated. In our case, it was simulated a server managing a certain region where there was an avatar whose client should receive state updates of the other avatars present in that region.
%In our case, it was simulated a server which should send state updates to the client responsible for one of the avatars in the region managed by that server.
Based on their locations and on the chosen interest management algorithm, the server decided which other avatars had a relevant state to the client in question.
